Fundamental Problems with the Evidence Base of Adolescent Depression Treatments

### 2024-04-09 Manuscript_for_rendering_in_word.qmd
Finished discussion referencing

Second, the low quality of control conditions of psychotherapy trials for adolescent depression should prompt a rethinking of how these studies are appraised and what needs to be done to create fair and realistic comparators. Indeed, investment of effort and funds should be directed into providing rigorous evidence that established depression psychotherapies are more efficacious than fair controls. There are examples of child psychological therapies, e.g. in social anxiety (CITE David Clark’s comparisons with behavioural interventions) where such rigor has been applied.

Third, our findings make clear the inherent difficulties of comparing psychotherapy with medication trials (CITE Sam and Andrea’s paper). The first obstacle is the comparability of the populations taking part. Even if a trial were designed to conduct a head-to-head comparison of psychotherapy with medication (as has been done CITE), it might sample the population of those who are indifferent to which one they would receive. And even in such a design, difficulties with blinding of the psychotherapy control would have to be overcome to draw valid inferences.

>>>>

Second, the low quality of control conditions of psychotherapy trials for adolescent depression should prompt a rethinking of how these studies are appraised and what needs to be done to create fair and realistic comparators. Indeed, investment of effort and funds should be directed into providing rigorous evidence that established depression psychotherapies are more efficacious than fair controls. There are examples of such psychotherapy RCTS, e.g. in social anxiety [@clarkCognitiveTherapyExposure2006a] where such rigor has been applied.

Third, our findings make clear the inherent difficulties of comparing psychotherapy with medication trials [@delgiovaneCombiningPharmacologicalNonpharmacological2019]. The first obstacle is the comparability of the populations taking part. Even if a trial were designed to conduct a head-to-head comparison of psychotherapy with medication (as has been done in [@FluoxetineCognitiveBehavioralTherapy2004]), it might sample the population of those who are indifferent to which one they would receive. And even in such a design, difficulties with blinding of the psychotherapy control would have to be overcome to draw valid inferences.
